{\bf {\Large Improving estimates of ecosystem metabolism by reducing effects of tidal advection on dissolved oxygen time series}}

### 2015-05-20 mat_proc.tex
nearly done with revisions

Tidal height is used as a proxy for advection because the model is empirical and only requires a variable that is correlated with the true measure of interest.  Moreover, tidal height measurements are more widely available and are generally monitored by water quality sondes.  An important distinction between tidal height and advection is that the two variables could provide different information about a tidal regime.  Tidal heights at the minimum and maximum of the range may be associated with periods of low advection when water masses are not moving rapidly past the sensor, whereas tidal heights near the mean may be more likely to have greater advection.  Accordingly, our use of tidal height should not be confused with a variable that is directly proportional to advection.  The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that the model can isolate.

Weighted regression was implemented as a moving window that allowed for estimation of \ac{DO} throughout the time series by adapting to variation through time as a function of tide. Regression models were estimated sequentially for each observation in the time series using dynamic weight vectors that change with the center of the window.  Weight vectors quantified the relevance of observations to the center of the window in respect to time, hour of the day, and tidal height.  Specifically, weights were assigned to each variable using a tri-cube weighting function \citep{Tukey77,Hirsch10}:
\begin{equation}
w= \left\{ 
  \begin{array}{l l}
    \left(1-\left(d/h\right)^3\right)^3 & \quad \textrm{if } |d| \leq h \\
    0 & \quad \textrm{if } |d| > h 
  \end{array} \right.
\end{equation}
where the weight $w$ of each observation is inversely proportional to the distance $d$ from the center of the window such that observations more similar to the point of reference are given higher weight in the regression.  Observations that exceed the maximum width of the window $h$ are assigned a weight of zero.  The tri-cube weighting function is similar to a Gaussian distribution such that weights decrease gradually from the center until the maximum window width is reached.  Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations from the center \citep{Hirsch10}.  The final weight vector for each observation is the product of the three separate weight vectors for time (day), hour, and tidal height. Windows for time and hour are used to weight observations based on distance (time) from the center of the window.  The window for tidal height is used to weight observations based on the difference from mean tide as a proportion of tide range.  For example, a half-window width of 0.5 means that observations are weighted proportionately within +/- 50\% of the tide range referenced to the tidal height in the center of the window. A low weight is given to an observation if any of the three weighting values were not similar to the center of the window since the final weight vector is the product of three weight vectors for each variable (see the link in the \hyperref[multi]{multimedia} section for graphical display of different weights).

>>>>

Tidal height is used as a proxy for advection because the model is empirical and only requires a variable that is correlated with the true measure of interest.  An important distinction between tidal height and advection is that the two variables could provide different information about a tidal regime.  Tidal heights at the minimum and maximum of the range may be associated with periods of low advection when water masses are not moving rapidly past the sensor, whereas tidal heights near the mean may be more likely to have greater advection.  Accordingly, our use of tidal height should not be confused with a variable that is directly proportional to advection.  The model only requires a variable that indicates a particular point in the tidal cycle, such that tidal height can be mapped to advection with quantifiable periodicity that the model can isolate.  Moreover, tidal height measurements are more widely available and are generally monitored by water quality sondes.

Weighted regression was implemented as a moving window that allowed for estimation of \ac{DO} throughout the time series by adapting to variation through time as a function of tide. Regression models were estimated sequentially for each observation in the time series using dynamic weight vectors that change with the center of the window.  Weight vectors quantified the relevance of observations to the center of the window in respect to time, hour of the day, and tidal height.  Specifically, weights were assigned to each variable using a tri-cube weighting function \citep{Tukey77,Hirsch10}:
\begin{equation}
w= \left\{ 
  \begin{array}{l l}
    \left(1-\left(d/h\right)^3\right)^3 & \quad \textrm{if } |d| \leq h \\
    0 & \quad \textrm{if } |d| > h 
  \end{array} \right.
\end{equation}
where the weight $w$ of each observation is inversely proportional to the distance $d$ from the center of the window such that observations more similar to the point of reference are given higher weight in the regression.  Observations that exceed the maximum width of the window $h$ are assigned a weight of zero.  The tri-cube weighting function is similar to a Gaussian distribution such that weights decrease gradually from the center until the maximum window width is reached.  Regressions that use simpler windows (e.g., boxcar approach) are  more sensitive to influential observations as they enter or leave the window, whereas the tri-cube function minimizes their effect through gradual weighting of observations from the center \citep{Hirsch10}.  The final weight vector for each observation is the product of the three separate weight vectors for time (day), hour, and tidal height (see the \ref{multi}). Windows for time (continuous throughout the day) and hour (within each day) are used to weight observations based on distance (time) from the center of the window.  The window for tidal height is used to weight observations based on the difference from mean tide as a proportion of tide range.  For example, a half-window width of 0.5 means that observations are weighted proportionately within +/- 50\% of the tide range referenced to the tidal height in the center of the window. A low weight is given to an observation if any of the three weighting values were not similar to the center of the window since the final weight vector is the product of three weight vectors for each variable (see the link in the \hyperref[multi]{multimedia} section for graphical display of different weights).

### 2015-05-20 multimedia.tex
nearly done with revisions

The supporting information for this manuscript includes a graphical illustration of the weighting scheme described in the material and procedures section (\href{https://beckmw.shinyapps.io/weights_widget}{https://beckmw.shinyapps.io/weights\_widget}), results for each simulation (\href{https://beckmw.shinyapps.io/detiding_sims/}{https://beckmw.shinyapps.io/detiding\_sims/}), and results for each case study (\href{https://beckmw.shinyapps.io/detiding_cases/}{https://beckmw.shinyapps.io/detiding\_cases/}).  Each link is a graphical summary of data based on interactive inputs to support the results in the manuscript.

A simple R package with a sample dataset and code to implement weighted regression is also available on GitHub, including functions to estimate ecosystem metabolism.  See the README file on the web page for download instructions and examples (\href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}).

>>>>

A simple R package with a sample dataset and code to implement weighted regression is also available on GitHub, including functions to estimate ecosystem metabolism.  See the README file on the web page for download instructions and examples (\href{https://github.com/fawda123/WtRegDO}{https://github.com/fawda123/WtRegDO}).  Interactive applications are also available that illustrate the weighting scheme described in the material and procedures section (\href{https://beckmw.shinyapps.io/weights_widget}{https://beckmw.shinyapps.io/weights\_widget}), results for each simulation (\href{https://beckmw.shinyapps.io/detiding_sims/}{https://beckmw.shinyapps.io/detiding\_sims/}), and results for each case study (\href{https://beckmw.shinyapps.io/detiding_cases/}{https://beckmw.shinyapps.io/detiding\_cases/}).  Each link is a graphical summary of data based on interactive inputs to support the results in the manuscript.

### 2015-05-20 swmp_noise.Rnw
nearly done with revisions

%%%%%%
% abstract
\newpage
\section{Abstract}
\input{abstract.tex}

@@

% example of error surfaces 
\centering\vspace*{\fill}
<<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\varepsilon_{pro}$), amount of observation observation uncertainty ($\\varepsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).  Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \\cref{fig:err_surf2}).">>=

@@

% example of error surfaces 
\centering\vspace*{\fill}
<<err_surf2, fig.height = 2.33, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.  Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \\cref{fig:err_surf1}).">>=

@@

# names for row categories in table
Parms <- c('$\\boldsymbol{DO_{die}}$', 
  '$\\boldsymbol{DO_{adv}}$',  
  '$\\boldsymbol{\\varepsilon_{pro}}$', 
  '$\\boldsymbol{\\varepsilon_{obs}}$'
  )
tab <- form_fun(to_tab[, -c(1:2)])
rows <- to_tab[, 2]
cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\varepsilon_{pro}$, $\\varepsilon_{obs}$).  Values represent averages from multiple simulations with common parameters.  For example, row one is a summary of all simulations for which the diel \\ac{DO} component was zero ($n=729$).'

@@

tab <- form_fun(to_tab[, -c(1:2)])
rows <- to_tab[, 2]
cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).  Values represent averages from multiple simulations with common window values For example, row one is a summary of all simulations for which the half window width was one day ($n=729$).'

>>>>

%%%%%%
% abstract
\newpage
\section{Abstract}
<<intro, child = 'abstract.Rnw'>>=
@

@@

% example of error surfaces 
\centering\vspace*{\fill}
<<err_surf1, fig.height = 3.5, fig.width = 5.5, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \\ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\\varepsilon_{pro}$), amount of observation observation uncertainty ($\\varepsilon_{obs}$), and strength of diel \\ac{DO} component ($DO_{die}$).  Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (as in \\cref{fig:err_surf2}).  See \\cref{tab:dtd_perf1} for a summary of results for each unique parameter.">>=

@@

% example of error surfaces 
\centering\vspace*{\fill}
<<err_surf2, fig.height = 2.33, fig.width = 6, cache = T, echo = F, eval = T, fig.cap = "Heat maps of correlations and errors (\\ac{RMSE}) for filtered \\ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \\ac{DO} ($DO_{bio}$) for varying half window widths: days, hour of day, and proportion of tidal range.  Each tile represents the correlation or error from results for a given combination of window widths averaged for all simulation parameters (as in \\cref{fig:err_surf1}).  See \\cref{tab:dtd_perf2} for a summary of results for each unique window.">>=

@@

# names for row categories in table
Parms <- c('$\\boldsymbol{DO_{die}}$', 
  '$\\boldsymbol{DO_{adv}}$',  
  '$\\boldsymbol{\\varepsilon_{pro}}$', 
  '$\\boldsymbol{\\varepsilon_{obs}}$'
  )
tab <- form_fun(to_tab[, -c(1:2)])
rows <- to_tab[, 2]
cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\\varepsilon_{pro}$, $\\varepsilon_{obs}$).  Values represent averages from multiple simulations with common parameters.  For example, row one is a summary of all simulations for which the diel \\ac{DO} component was zero ($n=729$).  See \\cref{fig:err_surf1} for results of all parameter combinations.'

@@

tab <- form_fun(to_tab[, -c(1:2)])
rows <- to_tab[, 2]
cap.val<-'Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \\ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).  Values represent averages from multiple simulations with common window values For example, row one is a summary of all simulations for which the half window width was one day ($n=729$). See \\cref{fig:err_surf2} for results of all window combinations.'
